Algebraic File Synchronization: Adequacy and Completeness

### 2018-07-09 p2/conclusion.tex
Spellcheck
Clarify update detector

A related problem is to extend the system and the proofs
to allow for cases where reconciliation cannot
complete fully
or if only a subset of the replicas are reconciled 
(e.g. due to network partitioning),
both of which would result in a state where different replicas
have different common ancestors, that is,
the updates specific to the replicas start from different points
in the update history of the filesystems.
Existing research can offer pointers as to how such cases can be modelled
in our algebraic system.
Parker et al. \cite{PPRS} and Cox and Josephson \cite{CJ}
describe version vectors (update histories) kept as metadata,
while Chong and Hamadi present distributed algorithms that allow incremental synchronization \cite{CH}.
Representing individual updates to files
in their modification histories (as described in \cite{CJ})
as separate commands could also enable an algebraic synchronizer to reconcile otherwise
conflicting updates and resolve partial reconciliations.

An extension of our results could also entail relaxing
the simplifying assumptions we used.
The model and algorithms could be extended to the $|\setd|>1$ case, % TODO write this
to enable the synchronizer
to identify and resolve a wider range of conflicts.
Also, pre- and post-processors could be designed that
reconcile meta-information in directories or encode them as files, % TODO write this
or present removing and creating the same content at a different node
not as separate updates but as a single rename (move).
These could contribute to the user-friendliness of the reconciler
and help avoid human error.

And finally, we hope that this work, together with \cite{NREC}, provides
a blueprint of constructing an algebra of commands for different storage protocols
(e.g. XML trees, mailbox folders, generic relational databases, etc.),
and of demonstrating the adequacy and completeness of update and conflict detection and reconciliation
algorithms defined over it.
This, in turn, can offer formal verification of the algorithms underlying
specific implementations in a variety of synchronizers.
Alternatively, by generalizing the parent--child relationships between filesystem nodes,
the demonstrated properties of minimal sequences of commands
and domains of sets of sequences ($\worksmeqsign$)
may also contribute to future research into algebraic structures
constrained by predefined sparse connections between their elements.

>>>>

A related problem is to extend the system and the proofs
to allow for cases where reconciliation cannot
complete fully
or if only a subset of the replicas are reconciled 
(e.g. due to network partitioning),
both of which would result in a state where different replicas
have different common ancestors, that is,
the updates specific to the replicas start from different points
in the update history of the filesystems.
Existing research can offer pointers as to how such cases can be modeled
in our algebraic system.
Parker et al. \cite{PPRS} and Cox and Josephson \cite{CJ}
describe version vectors (update histories) kept as metadata,
while Chong and Hamadi present distributed algorithms that allow incremental synchronization \cite{CH}.
Representing individual updates to files
in their modification histories (as described in \cite{CJ})
as separate commands could also enable an algebraic synchronizer to reconcile otherwise
conflicting updates and resolve partial reconciliations.

Future work could also investigate extending the 
model and algorithms to the $|\setd|>1$ case
so that directory metadata could be represented directly
as opposed to through pre- and post-processing as described
in \cref{sec_multidir}.

And finally, we hope that this work, together with \cite{NREC}, provides
a blueprint of constructing an algebra of commands for different storage protocols
(e.g. XML trees, mailbox folders, generic relational databases, etc.),
and of demonstrating the adequacy and completeness of the update and conflict detection and reconciliation
algorithms defined over it.
This, in turn, can offer formal verification of the algorithms underlying
specific implementations in a variety of synchronizers.
Alternatively, by generalizing the parent--child relationships between filesystem nodes,
the demonstrated properties of minimal sequences of commands
and domains of sets of sequences ($\worksmeqsign$)
may also contribute to future research into algebraic structures
constrained by predefined sparse connections between their elements.

### 2018-07-09 p2/ordering.tex
Clarify update detector

As we recognize that multiple orders may be possible,
let us describe our ordering algorithm by
defining a subset of the permutations, $\ordersetsign$.
The simplicity of the definition means that enumerating
all sequences in $\ordersetsign$ can be implemented by a straightforward
algorithm.

As we will see, $\ordersetsign$ contains all possible
sequences of commands that do not break all filesystems.
In \cref{simple_reorder_equiv}, we offer a formal proof of this.
The facts that all valid reorderings of simple sequences are equivalent,
and all equivalent sequences are reorderings are important 
properties of simple sequences.

\begin{mydef}[$\ordersetsign$]\mlabel{def_orderset}
For a simple set or sequence of commands $S$,
we use $\orderset{S}$ to denote the set of 
permutations of the commands in $S$
for which the following holds:
for any subsequence $\cxynv\cc\czwmv$,
\begin{itemize}
\item if $n=\parentf{m}$, then $[y]=[\vald]$
\item if $\parentf{n}=m$, then $y=\empt$.
\end{itemize}
\end{mydef}

As usual, by subsequence we mean a subset of the commands in $S$ in the same order as they are found in $S$.

@@

\begin{mylem}\mlabel{connected_changes}
Given a set of commands that
does not contain assertion commands,
and which can be applied to a filesystem in some order without breaking it,
and which contains commands on $n$ and $\nn$ where $\nn\descendant n$,
then the set must also contain a command
on each node between $n$ and $\nn$.
\end{mylem}
\begin{proof}
TODO: If something changes, everything changes up to the leaves.

Without loss of generality, we can assume that $\nn\neq\parent(n)$.
We prove that under the given conditions, the set must contain a command on $\parent(n)$.
Then, by reapplying this result, we know that the set must contain commands on every
ancestor of $n$ up to $\nn$.

@@

>>>>

As we recognize that multiple orders may be possible,
let us describe our ordering algorithm by
defining a subset of the permutations, $\ordersetsign$.
As we will see, $\ordersetsign$ contains all possible
sequences of commands that do not break all filesystems.
In \cref{simple_reorder_equiv}, we offer a formal proof of this.
The facts that all valid reorderings of simple sequences are equivalent,
and all equivalent sequences are reorderings, are important 
properties of simple sequences.

% Define this with the help or a relation / partial order? But then all sequences
% could be ordered while this definition rules out illegal ones.

\begin{mydef}[$\ordersetsign$]\mlabel{def_orderset}
For a simple set or sequence of commands $S$,
we use $\orderset{S}$ to denote the set of 
permutations of the commands in $S$
for which the following holds:
for any subsequence $\cxynv\cc\czwmv$,
\begin{itemize}
\item if $n=\parentf{m}$, then $y=\vald$,
\item if $\parentf{n}=m$, then $y=\empt$.
\end{itemize}
\end{mydef}

As usual, by subsequence we mean a subset of the commands in $S$ in the same order as they are found in $S$.
The simplicity of the definition means that sorting the commands
and enumerating
all sequences in $\ordersetsign$ can be implemented by a straightforward
algorithm.

@@

\begin{mylem}\mlabel{connected_changes}
Given a set of commands that
does not contain assertion commands,
and which can be applied to a filesystem in some order without breaking it,
and which contains commands on $n$ and $\nn$ where $\nn\descendant n$,
then the set must also contain a command
on each node between $n$ and $\nn$.
\end{mylem}
\begin{proof}
% If something changes, everything changes up to the leaves.
Without loss of generality, we can assume that $\nn\neq\parent(n)$.
We prove that under the given conditions, the set must contain a command on $\parent(n)$.
Then, by reapplying this result, we know that the set must contain commands on every
ancestor of $n$ up to $\nn$.

@@

Building on the results above, we finish by proving an important result,
namely, that command sequences can be made simple using syntactical operations.

\begin{mylem}\mlabel{can_simplify}
Each sequence of commands $\wkrsx{S}$
can be transformed into a simple sequence
$T\equiv S$ by applying repeatedly the following transformations:
{\rm(i)} swap neighboring commands on incomparable nodes;
{\rm(ii)} simplify neighboring commands on the same node;
{\rm(iii)} delete assertion commands.
\end{mylem}
\begin{proof}
We can remove all assertion commands, as based on
\cref{ax_assert} the resulting sequence extends the original one.
We therefore assume that $S$ contains no assertion commands.

\newcommand{\ucx}{\alpha}
\newcommand{\ucy}{\beta}
\newcommand{\ucz}{\gamma}

We aim to find a simple sequence that is equivalent to $S$.
If $S$ is already simple, there is nothing to do.
If it is not simple, let $\ucx$ be the first command
on a node $n$ that is already present in $S$,
and let $\ucy$ be the previous command on $n$.
Splitting $S$ around these commands we get
\[ S = S_0 \cc \ucy \cc S_1 \cc \ucx \cc S_2. \]

We now show that $S_1\indep\ucx$,
which is equivalent to $S_1\indep\ucy$ as both commands are on $n$.
We use an inverse proof and assume that there is a command on $m$ in $S_1$
where $m\descendant n$ or $m\ancestor n$.
We know both $\ucy\cc S_1$ and $S_1\cc\ucx$ are simple,
and both are defined on a filesystem
(on $S_0\aFS$ and $(S_0\cc\beta)\FS$, respectively).
Therefore from \cref{connected_changes} we know that
$S_1$ must also contain a command on $m'$ where
$m'=\parentf{n}$ or $n=\parentf{m'}$.
Let this command be $\ucz$.

If $n=\parentf{m'}$,
from \cref{simple_distant_pairs} 
and the relationship of nodes in construction and destruction pairs
we know that the subsequence
$\ucy\cc\ucz$ must be a construction pair,
and both $\ucy$ and $\ucz$ must be construction commands.
But we also know that $\ucz\cc\ucx$ must be a destruction pair,
and both $\ucz$ and $\ucx$ must be destruction commands.
This is a contradiction as a command cannot be both.
We arrive at the same contradiction if $m'=\parentf{n}$.

We therefore know that $S_1\indep\ucx$ and so
from \cref{ax_separate_commute}
\[ S \equiv S_0\cc\ucy\cc\ucx\cc S_1 \cc S_2. \]
We can apply \cref{ax_same_emptyseq,ax_same_singlec} to
$\ucy\cc\ucx$ and get a $S'\eqext S$ that contains one or two less commands on $n$.
Repeating the swap and the simplification we
can convert $S$ into an equivalent simple sequence $S^*$.
\end{proof}

### 2018-07-09 p2/upd-correct.tex
Minor improvements
Clarify update detector

Based on \cref{connected_changes} we can also show that
the update detector does function as intended as:

% TODO
TODO ENSURE UPDATE DETECTOR IS DEFINED.

@@

\begin{myth}\mlabel{update_works}
For a sequence of commands $U$ returned by the update detector
when comparing the non-broken $\FS^*$ to the original $\FS$,
$U\aFS = \FS^*$.
\end{myth}
\begin{proof}
Let $\FS$ be the original filesystem, and $\FS^*$
be the filesystem after the changes we intend to detect.
Let us assume that some mechanism recorded all changes that occurred
to $\FS$ until it reached the state $\FS^*$.
Our set of commands is sufficient to record any change, as
there are commands for every input and output type pair, therefore
we know that the changes can be recorded as a sequence of commands $S$
where $S\aFS=\FS^*$.
Note that $S$ will not contain assertion commands, as they do not
represent an actual change in the filesystem.

\newcommand{\ucx}{\alpha}
\newcommand{\ucy}{\beta}
\newcommand{\ucz}{\gamma}

We aim to find a simple sequence that is equivalent to $S$.
If $S$ is already simple, there is nothing to do.
If it is not simple, let $\ucx$ be the first command
on a node $n$ that is already present in $S$,
and let $\ucy$ be the previous command on $n$.
Splitting $S$ around these commands we get
\[ S = S_0 \cc \ucy \cc S_1 \cc \ucx \cc S_2. \]

We now show that $S_1\indep\ucx$,
which is equivalent to $S_1\indep\ucy$ as both commands are on $n$.
We use an inverse proof and assume that there is a command on $m$ in $S_1$
where $m\descendant n$ or $m\ancestor n$.
We know both $\ucy\cc S_1$ and $S_1\cc\ucx$ are simple,
and both are defined on a filesystem
(on $S_0\aFS$ and $(S_0\cc\beta)\FS$, respectively).
Therefore from \cref{connected_changes} we know that
$S_1$ must also contain a command on $m'$ where
$m'=\parentf{n}$ or $n=\parentf{m'}$.
Let this command be $\ucz$.

If $n=\parentf{m'}$,
from \cref{simple_distant_pairs} 
and the relationship of nodes in construction and destruction pairs
we know that the subsequence
$\ucy\cc\ucz$ must be a construction pair,
and both $\ucy$ and $\ucz$ must be construction commands.
But we also know that $\ucz\cc\ucx$ must be a destruction pair,
and both $\ucz$ and $\ucx$ must be destruction commands.
This is a contradiction as a command cannot be both.
We arrive at the same contradiction if $m'=\parentf{n}$.

We therefore know that $S_1\indep\ucx$ and so
from \cref{ax_separate_commute}
\[ S \equiv S_0\cc\ucy\cc\ucx\cc S_1 \cc S_2. \]
We can apply \cref{ax_same_emptyseq,ax_same_singlec} to
$\ucy\cc\ucx$ and get a $S'\eqext S$ that contains one or two less commands on $n$.
Repeating the swap and the simplification we
can convert $S$ into an equivalent simple sequence.
Let this sequence be $S^*$.

As $S^*\equiv S$, we know $S^*\aFS=\FS^*$, and that
$\FS^*$ and $\FS$ differ at exactly those nodes that $S^*$ has commands on.
If $U$ is the simple sequence returned by the update detector
when comparing $\FS^*$ to $\FS$, it must also contain commands on the same set of nodes,
and it must contain the same commands,
as for the command $\cxynv$, $[x]$ must be $[\FS(n)]$, and $y$ must be $\FS^*(n)$.
From this we know that $U$ is a permutation of $S^*$ and so $\orderset{U}=\orderset{S^*}$.

As the update detector uses $\ordersetsign$ to order the commands in $U$,
trivially $U\in\orderset{U}$,
and as $S^*$ is defined on $\FS$,
from \cref{order_is_only_possible} we know $S^* \in \orderset{S^*}$.
This means $U\in\orderset{S^*}$,
and so from \cref{simple_reorder_equiv} we know that $U\equiv S^*$,
from which $U\aFS=\FS^*$.
\end{proof}

>>>>

With $\ordersetsign$ we can now define our update detection algorithm.
Its inputs are the original filesystem $\FS$ which has been modified to yield $\FS_1$.
\begin{mydef}[Update detection]\mlabel{def_upddetect}
\begin{enumerate}
    \item For each node $n$ where the value in $\FS$ and $\FS_1$ differ, add the command $\cxynv$ to
        a set of commands $U$.
    \item % TODO
\end{enumerate}
\end{mydef}
% TODO from journal? Simplify sequence?

TODO ENSURE UPDATE DETECTOR IS DEFINED.

% There is a sequence between any two states: destroy one completely and construct the other one. (finite FS)

\begin{myth}\mlabel{update_works}
For a sequence of commands $U$ returned by the update detector
when comparing the non-broken $\FS^*$ to the original $\FS$,
$U\aFS = \FS^*$.
\end{myth}

@@

TODO And that it can be ordered?

\begin{proof}
Let $\FS$ be the original filesystem, and $\FS^*$
be the filesystem after the changes we intend to detect.
Let us assume that some mechanism recorded all changes that occurred
to $\FS$ until it reached the state $\FS^*$.
Our set of commands is sufficient to record any change, as
there are commands for every input and output type pair, therefore
we know that the changes can be recorded as a sequence of commands $S$
where $S\aFS=\FS^*$.
($S$ will not contain assertion commands, as they do not
represent an actual change in the filesystem.)
Based on \cref{can_simplify}, there is a simple sequence $S^*$
that is equivalent to $S$.

Therefore we know $S^*\aFS=\FS^*$, and that
$\FS^*$ and $\FS$ differ at exactly those nodes that $S^*$ has commands on.
If $U$ is the simple sequence returned by the update detector
when comparing $\FS^*$ to $\FS$, it must also contain commands on the same set of nodes,
and it must contain the same commands,
as for the command $\cxynv$, $[x]$ must be $[\FS(n)]$, and $y$ must be $\FS^*(n)$.
From this we know that $U$ is a permutation of $S^*$ and so $\orderset{U}=\orderset{S^*}$.

As the update detector uses $\ordersetsign$ to order the commands in $U$, trivially $U\in\orderset{U}$, % TODO
and as $S^*$ is defined on $\FS$,
from \cref{order_is_only_possible} we know $S^* \in \orderset{S^*}$.
This means $U\in\orderset{S^*}$,
and so from \cref{simple_reorder_equiv} we know that $U\equiv S^*$,
from which $U\aFS=\FS^*$.
\end{proof}

### 2018-07-09 p2/upd-def.tex
Clarify update detector

In a command-based reconciliation solution we assume that we have two sequences of commands
$A$ and $B$ that have modified a single filesystem $\FS$ 
yielding two different replicas $\FS_1$ and $\FS_2$ that we
need to reconcile. While it is conceivable that the sequences are based on a record of
all operations that modified the two filesystems, in most filesystem implementations
such records are not implemented, and therefore we must construct suitable sequences
by comparing $\FS_1$ (and $\FS_2$) to their common ancestor, $\FS$. 
This is called \emph{update detection.}

The set of commands $U$ necessary to transform $\FS$ into $\FS_1$ can be collected
by inspecting the union of non-empty nodes in the two filesystems.
If at node $n$, $\FS(n)=x$ and $\FS_1(n)=y\neq x$, then we add the command $\cxyaa{n}$ to $U$.
We can always do so as there is a command available for all combinations of input and output types and values.
We see that any suitable $U$ necessarily contains commands on all nodes at which the values have changed.

@@

% TODO Define, spell out update decetion algorithm

>>>>

In a command-based reconciliation solution we assume that we have two sequences of commands
$A$ and $B$ that have modified a single filesystem $\FS$ 
yielding two different replicas $\FS_1$ and $\FS_2$ which we
need to reconcile. While it is conceivable that the sequences are based on a record of
all operations that modified the two filesystems, in most filesystem implementations
such records are not implemented, and therefore we must construct suitable sequences
by comparing $\FS_1$ (and $\FS_2$) to their common ancestor, $\FS$. 
This is called \emph{update detection.}

The set of commands $U$ necessary to transform $\FS$ into $\FS_1$ can be collected
by inspecting the (necessarily finite) union of non-empty nodes in the two filesystems.
If at node $n$, $\FS(n)=x$ and $\FS_1(n)=y\neq x$, then we add the command $\cxyaa{n}$ to $U$.
We can always do so as there is a command available for all combinations of input and output types and values.
We see that any suitable $U$ necessarily contains commands on all nodes at which the values have changed.

@@
